---
title: "Capital Requirements for a Standardised IRB Model"
author: "Daniel Jordan"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Capital Requirements for a Standardised IRB Model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Internal Ratings Based Approach Models

IRB models can produce two outputs. In the case of Foundation IRB that output is a probability of default (PD). For Advanced IRB that is a probability of default and a loss given default (LGD). Using the PD and the LGD we can compute the capital requirement for a borrower. We can then compute the risk weighted assets of that borrower.

Steps to computing tier 1 capital:

  1. Compute 1-year PD  
  2. Compute capital requirement 
  3. Compute total RWA
  4. Compute Tier 1 capital
  
These are actually quite simple steps, but amazingly this process can take up to a year to complete.

### Computing Capital 

So now we have a dataset of borrowers, their exposure at each point in time and a PD. Given we are doing a foundation IRB model, we will use the same LGD for every borrower, prescribed by Basel as X. 

```{r}
lending_club %>% 
  mutate(K = capital_requirement(PD, LGD = ))
```

